```{r setup, echo = FALSE}
opts_chunk$set(fig.width=7, fig.height=5, comment = "", warning = FALSE, tidy = TRUE, highlight = TRUE, fig.path = 'figures/', dev = 'png', background = 'red', width = 100)
```
```{r libs}
library("trioClasses")
library("GWASTools")
library("CleftCNVAssoc")
```
```{r data, cache = FALSE, echo = FALSE}
data("beaty.scanAnnot")
data("beaty.snpAnnot")
data("fe")
intensfile <- NcdfIntensityReader("/media/sgy/lockdown/GWASTools/beaty.intensity.nc")
genofile <- NcdfGenotypeReader("/media/sgy/lockdown/GWASTools/beaty.geno.nc")
xyfile <- NcdfIntensityReader("/media/sgy/lockdown/GWASTools/beaty.intensity.xy.nc")
scan.ids <- as.character( getScanID( beaty_scanAnnot ) )
scan.names <- as.character( getVariable( beaty_scanAnnot, "scanName" ) )
names(scan.names) <- scan.ids
names(scan.ids) <- as.character(scan.names)
```

```{r ids, echo = FALSE}
fa.id <- as.character(completeTrios(fe.beaty)$fid)
ma.id <- as.character(completeTrios(fe.beaty)$mid)
names(fa.id) <- as.character(completeTrios(fe.beaty)$id)
names(ma.id) <- as.character(completeTrios(fe.beaty)$id)
```
First we create a vector of offspring IDs that we want plotted.
```{r foo, echo = TRUE}
offspring.vec <- as.character(completeTrios(fe.beaty)$id)
```
Now we incorporate it into a GRange object with the ranges in this case being the chr16 region for each of the offspring.
```{r grange}
gr <- GRanges( seqnames = rep('chr16',length(offspring.vec)), ranges = IRanges( start = 32404517, end = 32530051), id = offspring.vec )
```
Now we apply a function from CleftCNVAssoc to retrieve data from GWASTools objects.  These data exist only on an encrypted hard drive and enigma.
```{r raw, cache = FALSE}
raw.df.list <- getRaw( gr+1e6, intensfile = intensfile, snpAnnot = beaty_snpAnnot, scan.id = scan.ids, fa.id = fa.id, ma.id = ma.id, genofile = genofile, xyfile = xyfile )
```
Here are the data for the first offspring's first 5 markers.
```{r head}
head(raw.df.list[[1]],5)
```

```{r pos, echo = FALSE}
pos.vec <- raw.df.list[[1]]$pos
logr.list <- lapply(raw.df.list, function(df) df$logr )
logr.fa.list <- lapply(raw.df.list, function(df) df$logr.fa )
logr.ma.list <- lapply(raw.df.list, function(df) df$logr.ma )
```
Plot the logR values for everyone stratified by F,M,O.  Purple is offspring, red is father, and blue is mother.
```{r logrplot, fig.width = 24, echo = FALSE, cache = FALSE}
par(bg="lightyellow")
plot(1, type = 'n', xlim = range(pos.vec), ylim = c(-2,2) )
for( i in 1:length(logr.list) ){
  points(jitter(pos.vec,amount = 1e3), logr.list[[i]], pch = 20, col = "purple")
  }
plot(1, type = 'n', xlim = range(pos.vec), ylim = c(-2,2) )
for( i in 1:length(logr.fa.list) ){
  points(jitter(pos.vec,amount = 1e3), logr.fa.list[[i]], pch = 20, col = "red")
  }
plot(1, type = 'n', xlim = range(pos.vec), ylim = c(-2,2) )
for( i in 1:length(logr.ma.list) ){
  points(jitter(pos.vec,amount = 1e3), logr.ma.list[[i]], pch = 20, col = "blue")
  }
```
Not very informative so we turn to individual trios with an untransmitted deletion. First, we need to find a vector offspring IDs with an untransmitted deletion.  This is a property of the CNVMatrix within the FamilyExperiment object and can be manipulated with the non-exported method TrioAssay.  To begin we first subset the CNVMatrix on the chr16 region.
```{r subset}
chr16.gr <- GRanges( seqnames = 'chr16', ranges = IRanges( start = 32404517, end = 32530051))
(fe.beaty.chr16 <- fe.beaty[queryHits(findOverlaps(rowData(fe.beaty),chr16.gr))])
```
Now with the smaller FE object we can easily construct the trio-states.
```{r trios}
trioAssay.chr16 <- trioClasses:::TrioAssay(fe.beaty.chr16, type = "cnv")
trioStates.chr16 <- with(trioAssay.chr16, matrix( paste0(F,M,O), nrow = nrow(O), ncol = ncol(O)))
dimnames(trioStates.chr16) <- dimnames(trioAssay.chr16$O)
head(trioStates.chr16[,1:5],10)
```
Now we identify trio-cnv pairs with an untransmitted deletion, i.e., trio-states 100, 010, or 110.  (This is not a complete list of trio-states with a non-transmission.)
```{r untrans}
untrans.mat <- matrix(trioStates.chr16 %in% c("100","010","110"),nrow=nrow(trioStates.chr16),ncol=ncol(trioStates.chr16),byrow=FALSE,dimnames=dimnames(trioStates.chr16))
head(untrans.mat[,1:10],10)
```
And finally we find the IDs of those with more than zero untransmitted deletions.
```{r offspring}
offspring.chr16 <- rownames(untrans.mat)[which(rowSums(untrans.mat)>0)]
length(offspring.chr16)
head(offspring.chr16)
```
Now we have what we need to dig deeper into the cause of the under-transmission.
```{r logrplot2, fig.width = 24, echo = FALSE, cache = FALSE}
par(bg="lightyellow")
for( i in which(names(logr.list) %in% offspring.chr16) ){
  plot(1, type = 'n', xlim = range(pos.vec), ylim = c(-2,6) )
  points(pos.vec, logr.list[[i]], pch = 20, col = "purple")
  points(pos.vec, 4+logr.fa.list[[i]], pch = 20, col = "red")
  points(pos.vec, 4+logr.ma.list[[i]], pch = 20, col = "blue")
  foo <- reduce(rowData(fe.beaty.chr16)[untrans.mat[names(logr.list)[i],]])
  for( j in 1:length(foo) ) lines(c(start(foo[j]),end(foo[j])),rep(2,2), lwd = 2, col = "red")
  }
```
Now for the transmitted deletions in the chr16 region.
```{r trans}
trans.mat <- matrix(trioStates.chr16 %in% c("101","011","112"),nrow=nrow(trioStates.chr16),ncol=ncol(trioStates.chr16),byrow=FALSE,dimnames=dimnames(trioStates.chr16))
```
```{r offspring2}
offspring.chr16 <- rownames(trans.mat)[which(rowSums(trans.mat)>0)]
length(offspring.chr16)
head(offspring.chr16)
```
```{r logrplot3, fig.width = 24, echo = FALSE, cache = FALSE}
par(bg="lightyellow")
for( i in which(names(logr.list) %in% offspring.chr16) ){
  plot(1, type = 'n', xlim = range(pos.vec), ylim = c(-2,6) )
  points(pos.vec, logr.list[[i]], pch = 20, col = "purple")
  points(pos.vec, 4+logr.fa.list[[i]], pch = 20, col = "red")
  points(pos.vec, 4+logr.ma.list[[i]], pch = 20, col = "blue")
  foo <- reduce(rowData(fe.beaty.chr16)[trans.mat[names(logr.list)[i],]])
  for( j in 1:length(foo) ) lines(c(start(foo[j]),end(foo[j])),rep(2,2), lwd = 2, col = "purple")
  }
```
So we see an increase in density in the region that likely has an increase in false positive rate.  Is it the case that false positive rate and marker density are correlated?  Maybe in an effort to better identify known CNPs the manufacturers increased marker density which increased both false positive and negative rates.  And now, with trio data, we can identify the regions by under-transmission of deletions.

So, now we must find the marker density.  Let's find 